---
title: "Fast Knapsack Package"
author: "Anubhav Dikshit, Lennart Schilling, Thijs Quast" 
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This packages is a fast knapsack solver.
For a detailed background of the knapsack problem see [here](https://en.wikipedia.org/wiki/Knapsack problem).

Within this vignette, the example data.frame `knapsack_objects` is used to show how to work with the package.

It is a manually created data.frame with 2000 rows:
```{r, echo=FALSE, eval = TRUE}
set.seed(42)
n = 1000000
knapsack_objects <-
  data.frame(
    w=sample(1:4000, size = n, replace = TRUE),
    v=runif(n = n, 0, 10000)
  )
```
```{r, echo=TRUE, eval = TRUE}
head(knapsack_objects)
```

## Usage of the package
The package includes in total three functions which all return the result for a specified knapsack problem.

- `brute_force_knapsack(x, W)`
- `knapsack_dynamic(x, W)`
- `greedy_knapsack(x, W)`

For each function, inputs `x` and `W` have to be specified. 

`x` has to be a data.frame with two columns called `v` (for value) and `w` (for weight).
`W` is the specified maximum weight of the knapsack. 

All algorithms try to find - in a different way - the optimal filling of the knapsack to put in the maximum value considering the specified maximum weight `W`.

### `brute_force_knapsack(x, W)`
```{r, echo=FALSE, eval = TRUE}
library(utils)
library(rLab6)
library(knitr)
brute_force_knapsack <- function(x, W){
  original_value = x

    stopifnot(is.data.frame(x),is.numeric(W))
    stopifnot(W > 0)

  # reorder the items according to their weight to get near the maximum as soon as possible
  x <- x[rev(order(x[,1])),]

  # remove combinations that are invalid from the start
  # only consider items with a weight that is less than the capacity
  x <- x[x[,'w']<=W,]
  elements <- rownames(x)
    i=2
    optimum_value = 0
    selected_items = c()
    weights<-c()
    values<-c()
    while(i<=nrow(x))
    {
      w<-as.data.frame(combn(x[,1], i))
      v<-as.data.frame(combn(x[,2], i))

        sumw<-colSums(w) # most time consuming using profvis
        sumv<-colSums(v) # most time consuming using profvis

      weights<-which(sumw<=W)
      if(length(weights) != 0){
        values<-sumv[weights]
        optimum_value<-max(values)
        temp<-which((values)==optimum_value)
        maxValWghtIdx<-weights[temp]
        maxValWght<-w[, maxValWghtIdx]
        j<-1
        while (j<=i){
          selected_items[j]<-which(x[,1]==maxValWght[j])
          j=j+1
        }
      }
      i=i+1

    }
    elem <- subset(original_value, w %in% maxValWght)
    elem <- noquote(rownames(elem))

    return(list(value=round(optimum_value), elements=elem))
}
```
```{r, echo=TRUE, eval = TRUE}
brute_force_knapsack(x = knapsack_objects[1:8,], W = 3500)
```

### `knapsack_dynamic(x, W)`
```{r, echo=FALSE, eval = TRUE}
knapsack_dynamic <- function(x, W){

  stopifnot(is.data.frame(x),is.numeric(W))
  stopifnot(W > 0)

  # reorder the items according to their weight to get near the maximum as soon as possible
  x <- x[rev(order(x[,1])),]

  # remove combinations that are invalid from the start
  # only consider items with a weight that is less than the capacity
  x <- x[x[,'w']<=W,]
  elements <- rownames(x)
  w <-(x[,1])
  p <-(x[,2])

  n <- nrow(x)
  # initiate arrays
  x <- logical(n)
  F <- matrix(0, nrow = W + 1, ncol = n)
  G <- matrix(0, nrow = W + 1, ncol = 1)

  # forwarding part
  for (k in 1:n) {
    F[, k] <- G
    H <- c(numeric(w[k]), G[1:(W + 1 - w[k]), 1] + p[k])
    G <- pmax(G, H)
  }
  fmax <- G[W + 1, 1]

  # backtracking part
  f <- fmax
  j <- W + 1
  for (k in n:1) {
    if (F[j, k] < f) {
      x[k] <- TRUE
      j <- j - w[k]
      f <- F[j, k]
    }
  }

  inds <- which(x)
  elem <- elements[x]
  prof <- round(sum(p[inds]))
  elem <- noquote(elem)
  return(list(value = prof, elements = elem))
}
```
```{r, echo=TRUE, eval = TRUE}
knapsack_dynamic(x = knapsack_objects[1:8,], W = 3500)
```

### `greedy_knapsack(x, W)`
```{r, echo=FALSE, eval = TRUE}
library(Rcpp)
greedy_knapsack <- function(x, W, fast= NULL){

  stopifnot(is.data.frame(x),is.numeric(W))
  stopifnot(W > 0)

  x$v_by_w <- x$v/x$w

  # reorder the items according to their max profit per weight
  x <- x[rev(order(x[,3])),]

  x$max_weight <- W

  # remove combinations that are invalid from the start
  # only consider items with a weight that is less than the capacity
  x <- x[x[,'w']<=W,]
  elements <- rownames(x)
  x$running_weight <- cumsum(x$w)
  x$retain_in_bag <- ifelse(x$running_weight <= x$max_weight, "Retain", "Drop")
  x <- x[x$retain_in_bag == "Retain",]

  elem <- noquote(rownames(x))

  if(!is.null(fast)){
    prof <- round(vectorSum(x$v))
  }else{prof <- round(sum(x$v))}

  elem <- noquote(elem)


  return(list(value = prof, elements = elem))

}
```
```{r, echo=TRUE, eval = TRUE}
greedy_knapsack(x = knapsack_objects[1:8,], W = 3500)
```

## Speed of algorithms
The speed of the algorithms is tested. By doing so, the questions in the lab6 are answered.

### `brute_force_knapsack(x, W)`
Question: How much time does it takes to run the algorithm for n = 16 objects?
```{r, echo=TRUE, eval = TRUE}
start_time <- Sys.time()
brute_force_knapsack(x = knapsack_objects[1:16,], W = 3500)
end_time <- Sys.time()
end_time - start_time
```
### `knapsack_dynamic(x, W)`
Question: How much time does it takes to run the algorithm for n = 500 objects?
```{r, echo=TRUE, eval = TRUE}
start_time <- Sys.time()
knapsack_dynamic(x = knapsack_objects[1:500,], W = 3500)
end_time <- Sys.time()
end_time - start_time
```
### `greedy_knapsack(x, W)`
Question: How much time does it takes to run the algorithm for n = 1000000 objects?
```{r, echo=TRUE, eval = TRUE}
start_time <- Sys.time()
greedy_knapsack(x = knapsack_objects[1:1000000,], W = 3500)
end_time <- Sys.time()
end_time - start_time
```
### Code optimization

Question: What performance gain could you get by trying to improving your code?
```{r}
library(profvis)
library(microbenchmark)

profvis::profvis(greedy_knapsack(x = knapsack_objects[1:100000,], W = 3500))
```

Question: What performance gain could you get by using Rcpp and C++?
```{r eval=FALSE}
# Cant compile C function in vignette, thus run in console after loading the package
microbenchmark(greedy_knapsack(x = knapsack_objects[1:10000,], W = 3500),
               greedy_knapsack(x = knapsack_objects[1:10000,], W = 3500, fast = TRUE))
```

```{r, echo=FALSE}
result <- data.frame(expr =c("greedy_knapsack(x = knapsack_objects[1:10000, ], W = 3500)", "greedy_knapsack(x = knapsack_objects[1:10000, ], W = 3500, fast = TRUE)"), min = c(12.7947, 11.3946), lq =c(12.98865, 11.65790), mean =c(14.35950, 13.13924), median=c(13.85460, 12.59715))
```
Unit: milliseconds
```{r}
kable(result, format = "markdown")
```
Result: Thus the C code is about 1 millisecond or 7.69% faster than the R code

