---
title: "Linear Regression using Reference Class"
author: "Lennart Schilling, Thijs Quast, Anubhav Dikshit" 
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linear Regression using Reference Class}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package is an implementation of linear regression using linear algebra. 
Within this vignette, the example dataset `trees` is used. 

## Usage of the package
The package includes one function called `linreg(formula, data)` which returns a Reference Class object. 
```{r, echo=FALSE, eval = TRUE}
library(ggplot2)
linreg <- setRefClass("linreg",
                      fields = list(formula="formula",
                                    data="data.frame",
                                    m_X = "matrix",
                                    m_Y="matrix",
                                    Xt="matrix",
                                    XtX="matrix",
                                    betaestimates = "matrix",
                                    yfit="matrix",
                                    residual="matrix",
                                    nparameters = "integer",
                                    dof="integer",
                                    residualt = "matrix",
                                    residualvariance="numeric",
                                    residualstd = "numeric",
                                    betavariance = "matrix",
                                    bb = "numeric",
                                    tvalues="matrix",
                                    pvalues= "matrix",
                                    standardizedresiduals ="matrix",
                                    sqrtstresiduals = "matrix",
                                    export_formula = "formula",
                                    export_data = "character"
                                    ),
                      methods = list(
                        initialize = function(formula, data){
                          #Independent and dependent variables
                          m_X <<- model.matrix(formula, data)
                          m_Y <<- as.matrix(data[all.vars(formula)[1]])
                          # Transpose matrix X, and multiply + solve
                          Xt <<- t(m_X)
                          XtX <<- solve(Xt %*% m_X)
                          betaestimates <<- XtX %*% Xt %*% m_Y
                          # Estimate y
                          yfit <<- m_X %*% betaestimates
                          # Estimate residuals
                          residual <<- m_Y - yfit
                          # Determine degrees of freedom
                          nparameters <<- length(betaestimates)
                          dof <<- length(m_Y) - nparameters
                          # Variances
                          residualt <<- t(residual)
                          residualvariance <<- as.numeric((residualt %*% residual) / dof)
                          residualstd <<- sqrt(residualvariance)
                          betavariance <<- residualvariance * XtX
                          bb <<- diag(betavariance)
                          # t-values
                          tvalues <<- betaestimates/sqrt(bb)
                          # p-values
                          pvalues <<- 2 * pt(abs(tvalues), dof, lower.tail = FALSE)
                          # Standardized residuals for summary
                          standardizedresiduals <<- residual / sd(residual)
                          sqrtstresiduals <<- sqrt(abs(standardizedresiduals))

                          # saving names
                          export_formula <<- formula
                          export_data <<- deparse(substitute(data))
                        },
                        print = function(){
                          cat(paste("linreg(formula = ", format(export_formula), ", data = ", export_data , ")\n\n ", sep = ""))
                          setNames(round(betaestimates[1:nrow(betaestimates)],3),rownames(betaestimates))
                        },
                        resid = function(){
                          return(as.vector(residual))
                        },
                        pred = function(){
                          return(yfit)
                        },
                        coef = function(){
                          vec <- as.vector(betaestimates)
                          names(vec) <- colnames(m_X)
                          return(vec)
                        },
                        plot = function(){
                          plot1 <- ggplot(data.frame(yfit, residual), aes(y=residual, x=yfit)) + geom_point(shape=21, size=3, colour="black", fill="white")
                          plot1 <- plot1 + theme_linedraw() + theme(panel.grid.major = element_blank(),
                                                                            panel.grid.minor = element_blank(),
                                                                            plot.title = element_text(hjust = 0.5))
                          plot1 <- plot1 + stat_summary(fun.y=median, colour="red", geom="line", aes(group = 1))
                          plot1 <- plot1 + ggtitle("Residuals vs fitted") + xlab(paste("Fitted values \n lm(Petal.Length ~ Species)"))
                          plot2 <- ggplot(data.frame(yfit, sqrtstresiduals), aes(y=sqrtstresiduals, x=yfit)) + geom_point(alpha = 0.6, shape=21, size=3, colour="black", fill="white")
                          plot2 <- plot2 + theme_linedraw() + theme(panel.grid.major = element_blank(),
                                                                                 panel.grid.minor = element_blank(),
                                                                                 plot.title = element_text(hjust = 0.5))
                          plot2 <- plot2 + stat_summary(fun.y=median, colour="red", geom="line", aes(group = 1))
                          plot2 <- plot2 + ggtitle("Scale-Location") + xlab(paste("Fitted values \n lm(Petal.Length ~ Species)"))
                          plot2 <- plot2 + scale_x_continuous(breaks = seq(0.0, 1.5, by= 0.5))

                          plotlist <- list(plot1, plot2)
                          return(plotlist)
                        },
                        summary = function(){
                          "Prints the summary of linear regression model."
                          cat(paste("linreg(formula = ", format(export_formula), ", data = ", export_data, ") :\n\n ", sep = ""))
                          x <- setNames(as.data.frame(cbind(betaestimates,as.matrix(sqrt(bb)),tvalues, formatC(pvalues, format = "e", digits = 2), p_star_cal(pvalues))), c("Coefficients","Standard error","t values", "p values", ""))
                          myPrint(x)
                          cat(paste("\n\nResidual standard error: ", residualstd, " on ", dof, " degrees of freedom: ", sep = ""))
                        }
                      ))
myPrint = function(x, stripoff = FALSE) {
    print(x)
}
p_star_cal = function(p_value) {
  x <- ifelse(p_value > 0.1, " ",
              (ifelse(p_value > 0.05, " . ",
                      (ifelse(p_value > 0.01, "*",
                              (ifelse(p_value > 0.001, "**","***")))))))
  return(x)
}
```

```{r, echo=TRUE, eval=TRUE}
linregObject = linreg(Volume ~ ., trees)
```

To work with this object, several methods can be used:

- `print()`
- `resid()`
- `pred()`
- `coef()`
- `plot()`
- `summary()`

### print()
Returns the coefficients of the linear regression model.
```{r, echo=TRUE, eval=TRUE}
linregObject$print()
```

### resid()
Returns the vector of all resisduals.
```{r, echo=TRUE, eval=TRUE}
head(linregObject$resid())
```

### pred()
Returns the predicted values.
```{r, echo=TRUE, eval=TRUE}
head(linregObject$pred())
```

### coef()
Returns the coefficients as a named vector.
```{r, echo=TRUE, eval=TRUE}
linregObject$coef()
```

### plot()
Returns two plots related to the residuals.
```{r, echo=TRUE, eval=TRUE}
linregObject$plot()
```

### summary()
Returns a summary of the applied linregFunction.
```{r, echo=TRUE, eval=TRUE}
linregObject$summary()
```
